\vfill A Survey of Web Testing Frameworks and a Workflow for Testing Javascript Applications

### 2014-04-20 Paper/paper.tex
Simple notes to myself
Forgot to add notes about framework choices
Organizational changes, more content, a bit of everything...

\title{\vfill A Testing Framework for Web Applications} %\vfill gives us the black space at the top of the page
\author{
By Taggart Ashby \vspace{10pt} \\
}

@@

%%%%%%%%%%%%%%%%
%%% Abstract 
%%% VERY General overview of problem and solution in paper
%%%%%%%%%%%%%%%%
\begin{abstract}
\todo{Make this better}
We have brought together a number of testing tools from the web and created what we feel is a fully-featured, easy to use, testing framework for web applications.
\end{abstract}

@@

\subsection{Contributions of This Thesis}
- Framework
The goal of this thesis is to bring together tools that are already available for testing and combine them into a single framework that offers web application developers a single place to look for their testing needs. There are plenty of strong tools available and I hope to gather a subset of those tools to cover as much of testing spectrum as possible in an effort to save other developers time and energy in having to look at them all individually.
- Wade through all the tools currently out there
-- Focus in, no need to have every developer check out all 10000 tools
-- Weigh options

\subsection{Application Defaults \todo{This title sucks}}
- Node.js
- Browser based
- Javascript
- Database
- Open-source

\subsection{Tool Requirements \todo{Move this section elsewhere}}
- Free

@@

\subsection{Web Features}
Here I'll discuss some of the prevalent technologies that recently appeared in web applications and development. By no means is this an exhaustive list, just a sampling of the more prominent ones.

@@

% \subsubsection{JavaScript}
JavaScript, though not the only scripting language with web use, is the primary language used to compliment HTML when making web applications. It is an interpreted language that is most often associated with client-side functionality, but in recent times has extended to the server-side. JavaScript is a prototype based object oriented scripting language with dynamic typing and first-class functions. Much of JavaScript's popularity stems from this dynamic nature that allows for quick, and often dirty, implementation of features in an often iterative environment.

@@

\subsubsection{What is the Client-side Portion?}
First, I'll address what a client is. The simple answer is: You. Your phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service you're using to communicate with the web is the client.

@@

\subsubsection{What is the Server-side Portion?}
Once again, I'll outline a server and then delve into server-side programming. A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets, HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and a general number-crunching.

@@

% - CONSIDER REMOVING ALTOGETHER
% \subsubsection{Why Testing is Important}
% I'm consistently surprised by how few people take testing seriously when it comes to non-trivial applications. Many developers will leave testing to their users, i.e., wait for users to have problems and report them before developers will fix them. This is an issue for a number of reasons, a few of which are: frustrated customers are unlikely to continue using a buggy product and are dissuaded from buying future products from the same developer, certain bugs can be security loopholes that may allow users access to information they shouldn't have, in the worst cases a bug may cause damage to a person or property and the developers would be on the hook.

%%%%%%%%%%%%%%%%%%%%%%
%%% Implementation 
%%% 	Why I chose the tools I chose
%%% 	Scripts written to help out 
%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}

\section{Sample App and Testing Frameworks}

@@

\subsection{Sample ToDo App}
In order to test out the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our evaluation application. After some searching, I decided upon a Todo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} This application was selected because it included Node.js and had enough functionality that tests wouldn't be completely trivial while also not being overwhelming.
One of the challenges in testing that I quickly came across was the asynchronous nature of the MongoDB calls that were made in the sample app. Because of this headache, I will be discussing how the following frameworks deal with asynchronous test support as JavaScript is inherently asynchronous.
- Screenshots
- Example Navigation

\subsection{Unit Testing Frameworks}
The following frameworks are reviewed in chronological order of my evaluation. As such, comparisons will generally only be drawn between a framework and its predecessors in paper order. After all of the frameworks are outlined and evaluated, a longer section regarding the final decision will be made where comparisons will be drawn between all of them.

@@

\subsubsection{Jasmine \cite{Jasmine}}
The first of the unit testing frameworks that I looked at was Jasmine, a Behavior Driven Development (BDD) testing framework. BDD is based on Test Driven Development and adds some simplifications and patterns that attempt to bring both developers and businessmen into the software testing process. The idea behind Behavior Driven Development is that input from non-technical stakeholders as well as technical stakeholders can come together so that everyone understands what the project should do.

@@

\subsubsection{Mocha \cite{Mocha}}
Mocha is another BDD testing framework that is almost identical to Jasmine. Unlike Jasmine, however, Mocha is not completely self-contained and requires an additional assertion library of one's choosing. I will outline my choice of Chai.js below.

@@

Mocha's approach to asynchronous testing is quite a bit more straightforward than Jasmine's. In your ``it(...)'' anonymous function you pass an additional parameter called ``done'' which you then call when you know the asynchronous call has finished. This required me to edit the code I was testing to make all the async calls accept an additional callback, but it was a very small price to pay for extended testability. An example asynchronous test would look like this:
\begin{lstlisting}
describe("Asynch Mocha Test", function () {
  it("look like this", function (done) {
    myAsyncCall(param1, param2, function () {
      // This callback expected to run when async returns

@@

\paragraph{Assertion Library of Choice: Chai.js \cite{Chaijs}}
% Intentional space here
For the assertion library I chose Chai.js because it had the most flexibility for writing assertions that I'd found. It provides access to an Expect API, a Should API, and more general Assert API. The various APIs combined with Mocha's flexibilty led to a nice developer experience that can appeal to a variety of different developers.

\subsubsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the ``it'' functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}
An interesting feature that QUnit has that I've not seen elsewhere is the ability to specify how many assertions you expect to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let you know fewer assertions were called than expected.

@@

\subsubsection{Intern\cite{InternIO}}
Intern is a hybrid unit testing and browser testing framework that aims to be an all-in-one testing framework. Intern has unit testing, browser testing, and code coverage information all built into the same tool. Intern runs on its own server which give it access to the internals of your project, but caused problem with attempting to interact with a seperate server running the Node.js application. I was unable to get the browser testing portion of Intern to work because of the inability to set up cross-server communication easily and properly.

@@

Selenium's node incarnation is out-of-date and as such I had to change the selenium package to use the latest version of Selenium that works with the latest version of browsers. This required me to edit some of the javascript files that ran Selenium from the console using node. At the time of this writing, the Selenium node package was using Selenium 2.20 from over two years ago. The latest Selenium version is 2.40.

@@

\subsubsection{Jenkins \cite{Jenkins}}
Jenkins is the first of the continuous integration frameworks that I tried out. It was created by developers of a previous tool called Hudson when a dispute with Oracle led them to fork the Hudson project and rename it Jenkins. As of February 2014, the Jenkins organization has over 1,100 public repositories and 570 members. \cite{JenkinsGitHub} Jenkins has, according to their website, two main focuses: (1) building and testing software continuously and (2) monitoring executions of externally-run jobs. The first focus is the general goal for any CI framework. The second focus is of note and will be discussed after the initial evaluation.

@@

Getting a project running with Travis is fairly easy; you just create a yml file in your repository with certain lines pertaining to the language, needed services, commands to run, etc. I was able to set up everything I needed in six simple lines.

@@

From what I gathered, Travis appears to be targeted at companies, not individual developers. Travis requires you to host your projects on GitHub and only supports one process running at a time. However, Travis is incredibly easy to set up, does not require you to have your own server, comes with impressive built-in features, and works wonderfully if you are willing to open-source your repositories and put them on GitHub.

@@

\subsection{Misc Tools}
\subsection{Node Inspector}
Fantastic for debugging node.js server-side code!

@@

%%%%%%%%%%%%%%%%%%%%
%%% Related Work 
%%% Similar testing frameworks
%%% 	Why they're not as good as ours
%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
I was unable to find any works that attempted to bring existing technologies together to create a web testing framework. Instead, I came across a number of proprietary tools and research. For the proprietary tool an advantage of the framework I already laid out is that it is freely available and has support behind it for each individual component. There is no need to contact any paper authors to get software or support. I will, however, still outline them as they provided insight into what the problems in existing frameworks were, gaps in coverage, what was important in a framework, and inspiration and acknowledgment that this is a real problem that many people are trying to solve.

I also found research into common bugs, general testing techniques, and finding bugs that greatly improved my test cases and understanding.

@@

This 2-layer approach is interesting and provided me with some insight, but the system was made for PHP applications and I wasn't looking to implement a brand-new system and test PolyXpress with it all in one year.

@@

\subsubsection{``Continuous Testing with Ruby, Rails, and JavaScript'' \cite{BookContinuousTesting}}
The well known book by Rady and Coffin talks about Continuous Testing and has lots of examples. The last two chapters of the book, six and seven, talk about setting up a Continuous Testing environment for JavaScript using Node.js and a number of other tools. It fit almost perfectly with our project and was used as the basis of my testing setup. They use JSLint for linting, Jasmine for test cases, and Watchr for running the tests continuously. Lastly it discussing writing effective tests, which is always important.

@@

>>>>

\title{\vfill A Survey of Web Testing Frameworks and a Workflow for Testing Javascript Applications} %\vfill gives us the black space at the top of the page
\author{
By Taggart Ashby \vspace{10pt} \\
}

@@

%%%%%%%%%%%%%%%%
%%% Abstract 
%%% VERY General overview of problem and solution in paper
%%%%%%%%%%%%%%%%
\begin{abstract}
\todo{Make this better}
Web applications are quickly replacing standalone applications. As with any type of applicaiton, these web applications need to be tested to ensure proper functionality and reliability. There have been substantial efforts to create tools to help with the testing of web applicaitons but there is, as of yet, no standard or recommended workflow to insure speed of development and strength of application.

We have brought together a number of testing tools from the web and created what we feel is a fully-featured, easy to use, testing framework and workflow for javascript web applications.
\end{abstract}

@@

\subsection{Contributions of This Thesis}
- Framework
The goal of this thesis is to bring together tools that are already available for testing and combine them into a single framework that offers web application developers a single place to look for their testing needs. There are plenty of strong tools available and we hope to gather a subset of those tools to cover as much of testing spectrum as possible in an effort to save other developers time and energy in having to look at them all individually.
- Wade through all the tools currently out there
-- Focus in, no need to have every developer check out all 10000 tools
-- Weigh options

@@

\subsection{Web Features}
Here we'll discuss some of the prevalent technologies that recently appeared in web applications and development. By no means is this an exhaustive list, just a sampling of the more prominent ones.

@@

% \subsubsection{JavaScript}
JavaScript, though not the only scripting language with web use, is the primary language used to compliment HTML when making web applications. It is an interpreted language that is most often associated with client-side functionality, but in recent times has extended to the server-side. JavaScript is a prototype based object oriented scripting language with dynamic typing and first-class functions. Much of JavaScript's popularity stems from this dynamic nature that allows for quick, and often dirty, implementation of features in an iterative environment.

@@

\subsubsection{What is the Client-side Portion?}
First, we'll address what a client is. The simple answer is: You. Your phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service you're using to communicate with the web is the client.

@@

\subsubsection{What is the Server-side Portion?}
Once again, we'll outline a server and then delve into server-side programming. A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets, HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and a general number-crunching.

@@

% - CONSIDER REMOVING ALTOGETHER
% \subsubsection{Why Testing is Important}
% We're consistently surprised by how few people take testing seriously when it comes to non-trivial applications. Many developers will leave testing to their users, i.e., wait for users to have problems and report them before developers will fix them. This is an issue for a number of reasons, a few of which are: frustrated customers are unlikely to continue using a buggy product and are dissuaded from buying future products from the same developer, certain bugs can be security loopholes that may allow users access to information they shouldn't have, in the worst cases a bug may cause damage to a person or property and the developers would be on the hook.

%%%%%%%%%%%%%%%%%%%%%%
%%% Sample App 
%%%   Talk about test app
%%%%%%%%%%%%%%%%%%%%%%
\section{Sample App and Testing Frameworks}

\subsection{Sample ToDo App}
In order to test out the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our evaluation application. After some searching, We decided upon a Todo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} This application was selected because it included Node.js and had enough functionality that tests wouldn't be completely trivial while also not being overwhelming.
One of the challenges in testing that we quickly came across was the asynchronous nature of the MongoDB calls that were made in the sample app. Because of this headache, we will be discussing how the following frameworks deal with asynchronous test support as JavaScript is inherently asynchronous.
- Screenshots
- Example Navigation

%%%%%%%%%%%%%%%%%%%%%%
%%% Testing Frameworks
%%%   Talk about the various frameworks
%%%%%%%%%%%%%%%%%%%%%%
\section{Testing Frameworks}

@@

\subsubsection{The Average Web Application}
In an effort to create a good workflow and collection of tools, we have chosen to focus on web applications that uses Node.js, Javascript, Cloud Services, and a backing Database when searching for our tools. This choice was made because of the popularity of that kind of set up, as well as the fact that our evaluation application was set up in that way.

\todo{Need some sources and facts about Node.js popularity, average web application set up, etc. Can we back up our reasoning here?}

- Node.js
- Browser based
- Javascript
- Database
- Open-source

\subsubsection{Tool Requirements \todo{Move this section elsewhere}}
Our requirements for the selection of testing tools was relatively simple. In addition to meeting the requirements of our ``average web application'', the only other requirement was that the tool was free. What ``free'' means is that the developer would have access to at the least the minimum functionality needed to use the tool on \emph{one} application that is \emph{open-source}. What we quickly found is that web testing tools love open-source projects but are far less willing to provide free functionality to private or closed-source projects. Occasionally the cost of tools will be discussed but in general we didn't review tools that did not offer a free option.
- Free

\subsection{Unit Testing Frameworks}
The following frameworks are reviewed in chronological order of our evaluation. As such, comparisons will generally only be drawn between a framework and its predecessors in paper order. After all of the frameworks are outlined and evaluated, a longer section regarding the final decision will be made where comparisons will be drawn between all of them.

@@

\subsubsection{Jasmine \cite{Jasmine}}
The first of the unit testing frameworks that we looked at was Jasmine, a Behavior Driven Development (BDD) testing framework. BDD is based on Test Driven Development and adds some simplifications and patterns that attempt to bring both developers and businessmen into the software testing process. The idea behind Behavior Driven Development is that input from non-technical stakeholders as well as technical stakeholders can come together so that everyone understands what the project should do.

@@

\subsubsection{Mocha \cite{Mocha}}
Mocha is another BDD testing framework that is almost identical to Jasmine. Unlike Jasmine, however, Mocha is not completely self-contained and requires an additional assertion library of one's choosing. We will outline our choice of Chai.js below.

@@

Mocha's approach to asynchronous testing is quite a bit more straightforward than Jasmine's. In your ``it(...)'' anonymous function you pass an additional parameter called ``done'' which you then call when you know the asynchronous call has finished. This required us to edit the code we were testing to make all the async calls accept an additional callback, but it was a very small price to pay for extended testability. An example asynchronous test would look like this:
\begin{lstlisting}
describe("Asynch Mocha Test", function () {
  it("look like this", function (done) {
    myAsyncCall(param1, param2, function () {
      // This callback expected to run when async returns

@@

\paragraph{Assertion Library of Choice: Chai.js \cite{Chaijs}}
% Intentional space here
For the assertion library we chose Chai.js because it had the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and more general Assert API. The various APIs combined with Mocha's flexibilty led to a nice developer experience that can appeal to a variety of different developers.

\subsubsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the ``it'' functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions you expect to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let you know fewer assertions were called than expected.

@@

\subsubsection{Intern\cite{InternIO}}
Intern is a hybrid unit testing and browser testing framework that aims to be an all-in-one testing framework. Intern has unit testing, browser testing, and code coverage information all built into the same tool. Intern runs on its own server which give it access to the internals of your project, but caused problem with attempting to interact with a seperate server running the Node.js application. We were unable to get the browser testing portion of Intern to work because of the inability to set up cross-server communication easily and properly.

@@

Selenium's node incarnation is out-of-date and as such we had to change the selenium package to use the latest version of Selenium that works with the latest version of browsers. This required us to edit some of the javascript files that ran Selenium from the console using node. At the time of this writing, the Selenium node package was using Selenium 2.20 from over two years ago. The latest Selenium version is 2.40.

@@

\subsubsection{Jenkins \cite{Jenkins}}
Jenkins is the first of the continuous integration frameworks that we tried out. It was created by developers of a previous tool called Hudson when a dispute with Oracle led them to fork the Hudson project and rename it Jenkins. As of February 2014, the Jenkins organization has over 1,100 public repositories and 570 members. \cite{JenkinsGitHub} Jenkins has, according to their website, two main focuses: (1) building and testing software continuously and (2) monitoring executions of externally-run jobs. The first focus is the general goal for any CI framework. The second focus is of note and will be discussed after the initial evaluation.

@@

Getting a project running with Travis is fairly easy; you just create a yml file in your repository with certain lines pertaining to the language, needed services, commands to run, etc. We were able to set up everything we needed in six simple lines.

@@

From what we gathered, Travis appears to be targeted at companies, not individual developers. Travis requires you to host your projects on GitHub and only supports one process running at a time. However, Travis is incredibly easy to set up, does not require you to have your own server, comes with impressive built-in features, and works wonderfully if you are willing to open-source your repositories and put them on GitHub.

@@

\subsubsection{Grunt.js}
- Task Runner
- Watch functionality
- Easy setup

\subsection{Misc Tools}
\subsubsection{Node Inspector}
Fantastic for debugging node.js server-side code!

%%%%%%%%%%%%%%%%%%%%%%
%%% Implementation 
%%%   Why we chose the tools we chose
%%%   Scripts written to help out 
%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}

\subsection{Chosen Toolset}

- Unit Testing: Mocha w/ Chai
- Coverage: Istanbul
- CI: MagnumCI
- Browser Testing: Nightwatch
- Task Runner: Grunt
- Package Manager: NPM, Bower

\subsection{Setup}

- Note OS X 'ulimit' issue with Grunt-watch

\subsection{Workflow}
- Unit Testing on save
- CI used for running browser tests

- 'Manually' instantiate Sauce Labs?

@@

\subsection{PolyXpress (PX)}

\subsubsection{PolyXpress Author}
Our tests covered the Authoring portion of PX.

\subsection{Our Setup}
- 15-inch MacBook Pro (Late 2008 model)
-- Processor  2.53 GHz Intel Core 2 Duo
-- Memory  4 GB 1067 MHz DDR3
-- Graphics  NVIDIA GeForce 9400M 256 MB
--- Not a particularly powerful computer, just your average everyday laptop.

- OS 10.9.2
-- Terminal
- Sublime w/ plugins
- Netbeans

%%%%%%%%%%%%%%%%%%%%
%%% Related Work 
%%% Similar testing frameworks
%%% 	Why they're not as good as ours
%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
We were unable to find any works that attempted to bring existing technologies together to create a web testing framework. Instead, we came across a number of proprietary tools and research. Compared to these proprietary tools, the framework we already laid out is free and each part is supported by a community. There is no need to contact any paper authors to get software or support. We will, however, still outline them as they provided insight into what the problems in existing frameworks were, gaps in coverage, what was important in a framework, and inspiration and acknowledgment that this is a real problem that many people are trying to solve.

We also found research into common bugs, general testing techniques, and finding bugs that greatly improved our test cases and understanding.

@@

This 2-layer approach is interesting and provided us with some insight, but the system was made for PHP applications and we weren't looking to implement a brand-new system and test PolyXpress with it all in one year.

@@

\subsubsection{``Continuous Testing with Ruby, Rails, and JavaScript'' \cite{BookContinuousTesting}}
The well known book by Rady and Coffin talks about Continuous Testing and has lots of examples. The last two chapters of the book, six and seven, talk about setting up a Continuous Testing environment for JavaScript using Node.js and a number of other tools. It fit almost perfectly with our project and was used as the basis of our testing setup. They use JSLint for linting, Jasmine for test cases, and Watchr for running the tests continuously. Lastly it discussed writing effective tests, which is always important.

@@

-- Extension to all of PX?
-- Extension to non-proprietary open-source project
